---
title: "Volunteer Report Cité-Unis"
shorttitle: "Cité-Unis"
author:
  - name: Jan Pfänder
    corresponding: true
    email: janlukas.pfaender@gmail.com
    affiliations: Paris
    
abstract: "This report analyzes survey data for four different cohorts of Cité-Unis volunteers who did their service civique (2020-2024)."

mask: true

format:
  html: default
  apaquarto-docx: default  
  apaquarto-pdf:
    # Can be jou (journal), man (manuscript), stu (student), or doc (document)
    documentmode: doc
    toc: true
    toc-depth: 4
    # header-includes:
    #   - \usepackage{float}
    #   - \floatplacement{table}{H}

always_allow_html: true
---

```{r}
library(tidyverse)
library(readxl)
library(labelled) # for extracting labels from labelled data
library(sjlabelled)
library(ggalluvial)  # For alluvial (sankey-style) plots
library(sf)         # For GIS magic
library(rmapshaper) # simplify maps (so that they're less complex and take less time to handle)
library(wesanderson)
library(kableExtra)
library(broom)
library(gtsummary)
library(flextable)
library(gt)
library(ggpubr)
library(DescTools)
#library(brms)
```

```{r functions}
# load functions
source("functions/functions.R")
```

```{r}
# read data
load("data/cleaned_promo_combined.RData")
combined <- combined_data

# read map
map <- readRDS("data/map.rds")

# read codebook
codebook <- read_csv("data/codebook.csv") |> 
  # display answer options more nicely
  mutate(answer_options = str_replace_all(answer_options, "\\n", "; ")) 

# define by hand demographic variables (for tables and computations)
demographic_variables <- c(
  "type_volontaire", "sexe",
  "refugie", "age_categorie", "zone_residence",
  "niveau_etudes", "handicap_auto_declare", "handicap_administratif",
  "duree_prevue_sc_mois", "duree_reelle_sc_mois", "rupture"
)

# define other demographic variables that are not reported in tables, but used for subsetting data sets
demographic_variables_not_reported <- c("id_jeune", "region", "site", "type_volontaire", "decrocheur", "nationalite", "neets", "programme_1", "programme_2", "rupture_negative", "rupture_valence")
```


```{r}
# define colors
program_colours <- c(
  "#2E4057",  # deep blue
  "#A6332A",  # earthy red
  "#2F6F4E",  # forest green
  "#B67C44",  # warm amber/brown
  "#5D4E7A",  # muted purple
  "#E69F3D",  # golden orange
  "#4E9F9F"   # teal/blue-green
)
```

\clearpage

# Introduction

This report analyzes survey data for four different cohorts of Cité Unis volunteers who did their service civique in France between 2020 and 2024.

@sec-who provides information on who the volunteers are, in terms of age, sex, level of education and geographical location. @sec-attrition provides an overview on how many volunteers answered to questionnaires at different stages of their service civique (every volunteer is asked to fill out three questionnaires in total).

In @sec-within-change, we look at how volunteers have changed certain attitudes and views during their service civique, and which demographic and other factors are associated with changes.

We then look at several outcomes of interest in detail, namely whether volunteers end their contract early (@sec-rupture), how satisfied they are (@sec-satisfaction), and how confident they are about their future (@sec-perception-avenir). For these outcomes, we analyze whether there are trends across the different cohorts, and whether the outcomes are associated with demographic and other factors. 

In @sec-programs we investigate demographic differences between different key programs offered by Cité Unis for their volunteers. 

The report ends with a brief conclusion on methodological notes for future questionnaires. 

For these analyses, we rely on questionnaires collected by Cité Unis for four different cohorts of volunteers who did their service civique for a year (2020-2021; 2021-2022; 2022-2023; 2023-2024). These questionnaires are very extensive. For the present analyses, we selected a subset of key questions (a full list can be found in the [codebook](codebook.qmd))^[Note that the questions have been selected based on the questionnaire of the first cohort (2020/21). As a result, potentially interesting variables that only appear in later questionnaires will not appear here.]. 

Before diving into the results, a note on caution in interpreting the results presented in is report: Whenever we speak of "predictions" that simply means statistical associations---mere observations of differences between groups. This report does **not** provide any evidence that would warrant causal conclusions---answers as to **why** we observe these differences. 

Because the data underlying this report is observational and---for most of it---does not allow to track opinion change (because most questions have only been asked once per cohort), the main objective of this report is not to provide specific and definitive answers. Instead, the aim of this report is to help Cité Unis in identifying patterns and sparking question that may inform new evaluation procedures. The quantitative results presented here can act as pointers, but they are not problem diagnoses. They always need to be accompanied by qualitative assessments of *why* we observe certain differences, and what could or could not be done about them. 

# Who are the volunteers ? {#sec-who}

In this section, we review some demographic variables in detail. An extensive summary table with sample demographics across the different cohorts can be found in [Tables](tables.qmd). 

## Geographic location

```{r descriptives-demographics}
# get average participants per year for departments
participants_per_year_department <- map |> 
  st_drop_geometry() |> 
  group_by(NOM) |> 
  summarise(n = round(mean(n, na.rm = TRUE))) |> 
  arrange(desc(n))

n_departments <- sum(!is.na(participants_per_year_department$n))

# trend data by departments
trend_department <- map |> 
  st_drop_geometry() |> 
  pivot_wider(names_from = promo, 
              values_from = n) |> 
  mutate(trend = `23-24` - `20-21`) |> 
  summarize(
    mean_trend = mean(trend, na.rm = TRUE), 
    trend_positive = sum(trend > 0, na.rm = TRUE), 
    trend_negative = sum(trend < 0, na.rm = TRUE)
  ) |> 
  rounded_numbers()

# average trends for promo
n_per_promo <- combined |>
  group_by(promo) |>
  summarise(n = n_distinct(id_jeune)) |>
  # Extract starting year of each promo
  mutate(year = as.numeric(str_sub(promo, 1, 2))) |>  
  arrange(year) 
  
trend_per_year <- lm(n ~ year, data = n_per_promo) |> 
  tidy() %>%
  split(.$term)

```

Volunteers came from `r n_departments` different departments (see @fig-map-repartition). On average, across the different cohorts, most volunteers came from `r participants_per_year_department$NOM[1]` (n = `r participants_per_year_department$n[1]`), followed by `r participants_per_year_department$NOM[2]` (n = `r participants_per_year_department$n[2]`) and `r participants_per_year_department$NOM[3]` (n = `r participants_per_year_department$n[3]`). 

Overall, there has been a steady increase in volunteers, from `r first(n_per_promo$n)` in `r first(n_per_promo$promo)` to `r last(n_per_promo$n)` in `r last(n_per_promo$promo)`, and an average increase of `r trend_per_year$year$estimate` per year. Since the `r first(n_per_promo$promo)` cohort, each departement has on average increased by `r trend_department$mean_trend`. There were `r trend_department$trend_positive` departements who saw an increase, and `r trend_department$trend_negative` who saw a decrease (see @fig-map-trend). For details on the trend of each departement, see [Tables](tables.qmd). 

```{r}
#| label: fig-map-repartition
#| fig-cap: Répartition des volontaires en France à travers le temps. 

pal <- wes_palette("Zissou1", 100, type = "continuous")

ggplot() +
  geom_sf(data = map, aes(fill = n)) +
  scale_fill_gradientn(colours = pal, na.value = "white") + 
  theme_minimal() +
  coord_sf(datum = NA) +  # removes the coordinate axes and graticules
  labs(
    title = "Répartition des volontaires en France",
    fill = "Nombre de volontaires"
  ) +
  facet_wrap(~promo)
```

```{r}
#| label: fig-map-trend
#| fig-cap: Evolution de recrutement pour la promo de 2023-24 par rapport à 2020-21. 
#| 
# make a trend map of volunteers
map |> 
  pivot_wider(names_from = promo, 
              values_from = n) |> 
  mutate(trend = `23-24` - `20-21`) |> 
ggplot() +
  geom_sf(aes(fill = trend)) +
  scale_fill_gradient2(low = "#2C7BB6", mid = "grey", high = "#D7191C",
                           midpoint = 0, na.value = "white") +
  theme_minimal() +
  coord_sf(datum = NA) +  # removes the coordinate axes and graticules
  labs(
    title = "Difference Récrutement entre 2023 et 2020",
    fill = "Nombre de volontaires"
  )
```

## Age

As shown in @fig-age, across all promos, the majority of volunteers is between 18 and 20 years old. The share of the age group of 16-17 has been increasing rapidly, roughly doubling from 13% in 2020 to 24% in 2024.

```{r}
#| label: fig-age
#| fig-cap: "Number of volunteers per age group, within the different promos. Note that, in the percentage plot, the percentages are relative to all volunteers from the respective promo."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# Prepare data: Count unique participants per source
age_data <- combined |> 
  group_by(promo, age_categorie) |> 
  summarise(n_participants = n_distinct(id_jeune)) |> 
  group_by(promo) |> 
  mutate(share = n_participants/sum(n_participants)) |> 
  ungroup()


# Plot attrition trend across sources
ggplot(age_data, aes(x = age_categorie, y = n_participants)) +
geom_col() +
  geom_text(aes(label = n_participants), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  theme_minimal() +
  labs(
    title = "Age groups",
    x = "Age group",
    y = "Number of Volunteers"
  ) + 
  plot_theme +
  facet_wrap(~promo)

# PERCENTAGES

color <- wes_palette("BottleRocket2")[3]

ggplot(age_data,
       aes(x = age_categorie, y = share)) +
  geom_col(fill = color) +
  geom_text(aes(label = paste0(round(share*100, 1), "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Age groups",
    x = "Age group",
    y = "Share of volunteers"
  ) +
  plot_theme +
  facet_wrap(~promo)
```

## Education

@fig-education shows that volunteers with a "Bac + 3 et plus" are relatively rare, compared to the other education levels. The share of volunteers with a "Bac à Bac + 2" has been constantly decreasing, from 48% in 2020-21 to 40% in 2023-24. By contrast, volunteers "Infra-bac", have been increasing from 32% in 2020-21 to 43% in 2023-24. 

```{r}
#| label: fig-education
#| fig-cap: "Number of volunteers per education level, within the different promos. Note that in the percentage plot, the percentages are relative to all volunteers from the respective promo."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# Prepare data: Count unique participants per source
education_data <- combined |> 
  group_by(promo, niveau_etudes) |> 
  summarise(n_participants = n_distinct(id_jeune)) |> 
  group_by(promo) |> 
  mutate(share = n_participants/sum(n_participants)) |> 
  ungroup()


# Plot attrition trend across sources
ggplot(education_data, aes(x = niveau_etudes, y = n_participants)) +
geom_col() +
  geom_text(aes(label = n_participants), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  theme_minimal() +
  labs(
    title = "Education",
    x = "Level of education",
    y = "Number of volunteers"
  ) + 
  plot_theme +
  facet_wrap(~promo)

# PERCENTAGES

color <- wes_palette("BottleRocket2")[4]

ggplot(education_data,
       aes(x = niveau_etudes, y = share)) +
  geom_col(fill = color) +
  geom_text(aes(label = paste0(round(share*100, 1), "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Education",
    x = "Level of education",
    y = "Share of volunteers"
  ) +
  plot_theme +
  facet_wrap(~promo)
```

## Sex

On average, more women (~60%) are being volunteers than men. This difference is relatively stable across cohorts---if anything, the data suggests a slight trend towards an increasing share of women over time (@fig-sex). 

```{r}
#| label: fig-sex
#| fig-cap: "Number of male and female volunteers, within the different promos. Note that in the percentage plot, the percentages are relative to all volunteers from the respective promo."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# Prepare data: Count unique participants per source
sex_data <- combined |> 
  group_by(promo, sexe) |> 
  summarise(n_participants = n_distinct(id_jeune)) |> 
  group_by(promo) |> 
  mutate(share = n_participants/sum(n_participants)) |> 
  ungroup()


# Plot attrition trend across sources
ggplot(sex_data, aes(x = sexe, y = n_participants)) +
geom_col() +
  geom_text(aes(label = n_participants), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(3, 5)]) +
  theme_minimal() +
  labs(
    title = "Sex",
    x = "Sex",
    y = "Number of volunteers"
  ) + 
  guides(fill = FALSE) +
  plot_theme +
  facet_wrap(~promo)

# PERCENTAGES
ggplot(sex_data,
       aes(x = sexe, y = share, fill = sexe)) +
  geom_col() +
  geom_text(aes(label = paste0(round(share*100, 1), "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(3, 5)]) +
  labs(
    title = "Sex",
    x = "Sex",
    y = "Share of volunteers"
  ) +
  guides(fill = FALSE) +
  plot_theme +
  facet_wrap(~promo)
```

# Attrition {#sec-attrition}

As part of their service civique, each volunteer is asked to fill out three different questionnaires: the first one at the onset of their service civique (later on referred to as "q1"), the second one at the end ("q2"), and the third one six months after their service civique ("q1"). Answering these questionnaires is voluntary, and response rates typically decline considerably from the first to the third questionnaire. These declining response rates are referred to as "attrition". Some volunteers do not even respond to any one questionnaire^[Cité Unis nevertheless has some administrative data on volunteers who did not respond to a single questionnaire, for example which program they have been in. In some of the plots in this report, notably in @sec-attrition, these volunteers appear under the label "q0".]. 

Attrition can can have many causes. Here we distinguish between two explanations: ruptures, i.e. volunteers ending their service civique early (@sec-rupture) and survey fatigue, i.e. volunteers who continue their service civique but do not fill out the questionnaires (@fig-attrition-norupture). 

@fig-attrition shows attrition generally, while @fig-attrition-norupture excludes ruptures, thereby giving an estimate of survey fatigue. This latter figure shows that survey fatigue is an issue, but also suggests that there is a slightly positive trend towards reduced survey fatigue: more and more volunteers seem to answering the questionnaires at least for the first two questionnaires (q1 and q2), but less so for the third questionnaire (q3). 

```{r}
#| label: fig-attrition
#| fig-cap: "Number of volunteers per survey time point. Volunteers who appear under 'q0' have participated in the program but have not even filled out the first questionnaire (q1). Note that in the percentage plot, the percentages are relative to all volunteers from the respective promo. This plot **includes** ruptures."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# Prepare data: Count unique participants per source
attrition_data <- combined |> 
  group_by(promo, source) |> 
  summarise(n_participants = n_distinct(id_jeune)) |> 
  # remove the programmes (since it is not a survey time point)
  filter(source != "programmes") 


# Plot attrition trend across sources
ggplot(attrition_data, aes(x = source, y = n_participants)) +
geom_col() +
  geom_text(aes(label = n_participants), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  theme_minimal() +
  labs(
    title = "Attrition",
    x = "Time Point",
    y = "Number of Participants"
  ) + 
  plot_theme +
  facet_wrap(~promo)

# PERCENTAGES

# Step 1: Get baseline (q0 + q1) counts per promo
baseline_counts <- combined |>
  filter(source == "q0" | source == "q1") |>
  group_by(promo) |>
  summarise(baseline_n = n_distinct(id_jeune), .groups = "drop")

# Step 2: Count participants per source and promo
attrition_percent <- combined |>
  group_by(promo, source) |>
  summarise(n_participants = n_distinct(id_jeune), .groups = "drop") |>
  filter(source != "programmes") |> 
  left_join(baseline_counts, by = "promo") |>
  mutate(percent = n_participants / baseline_n)

# Step 3: Plot percentages relative to q0 and q1 

color <- wes_palette("Moonrise2")[1]

ggplot(attrition_percent |> 
         # remove q0 baseline
         filter(source != "q0"), 
       aes(x = source, y = percent)) +
  geom_col(fill = color) +
  geom_text(aes(label = paste0(round(percent*100, 1), "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Attrition",
    x = "Time Point",
    y = "Share of participants"
  ) +
  plot_theme +
  facet_wrap(~promo)
```

```{r}
#| label: fig-attrition-norupture
#| fig-cap: "Number of volunteers per survey time point,  **excluding** ruptures. Volunteers who appear under 'q0' have participated in the program but have not even filled out the first questionnaire (q1). Note that in the percentage plot, the percentages are relative to all volunteers from the respective promo."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# Prepare data: Count unique participants per source
attrition_data <- combined |> 
  filter(rupture == "no_rupture") |> 
  group_by(promo, source) |> 
  summarise(n_participants = n_distinct(id_jeune)) |> 
  # remove the programmes (since it is not a survey time point)
  filter(source != "programmes") 


# Plot attrition trend across sources
ggplot(attrition_data, aes(x = source, y = n_participants)) +
geom_col() +
  geom_text(aes(label = n_participants), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  theme_minimal() +
  labs(
    title = "Survey fatigue (ruptures excluded)",
    x = "Time Point",
    y = "Number of Participants"
  ) + 
  plot_theme +
  facet_wrap(~promo)

# PERCENTAGES

# Step 1: Get baseline (q0 + q1) counts per promo
baseline_counts <- combined |>
  filter(rupture == "no_rupture") |> 
  filter(source == "q0" | source == "q1") |>
  group_by(promo) |>
  summarise(baseline_n = n_distinct(id_jeune), .groups = "drop")

# Step 2: Count participants per source and promo
attrition_percent <- combined |>
  filter(rupture == "no_rupture") |> 
  group_by(promo, source) |>
  summarise(n_participants = n_distinct(id_jeune), .groups = "drop") |>
  filter(source != "programmes") |> 
  left_join(baseline_counts, by = "promo") |>
  mutate(percent = n_participants / baseline_n)

# Step 3: Plot percentages relative to q0 and q1 

color <- wes_palette("Moonrise2")[3]

ggplot(attrition_percent |> 
         # remove q0 baseline
         filter(source != "q0"), 
       aes(x = source, y = percent)) +
  geom_col(fill = color) +
  geom_text(aes(label = paste0(round(percent*100, 1), "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  labs(
    title = "Survey fatigue (ruptures excluded)",
    x = "Time Point",
    y = "Share of participants"
  ) +
  plot_theme +
  facet_wrap(~promo)
```

# How have volunteers changed their attitudes? {#sec-within-change}

In the selection of variables made for this report, there are only two questions that volunteers have been asked at several time points during their service civique: 

1. "Avez-vous voté lors des dernières élections (locales ou nationales) ?" (@sec-vote)

2. "En général, pensez-vous que votre action individuelle peut contribuer à changer la société ?" (@sec-individual-action). 

## Voting {#sec-vote}

In this section, we look at how volunteers have changed regarding their voting behavior. Voting behavior was measured with the question: "Avez-vous voté lors des dernières élections (locales ou nationales) ?", which has been asked at time points q1 and q2. @fig-vote_derniers_elections-overview shows that, averaged across all cohorts and time points, the majority of volunteers reports having voted.

```{r}
#| label: fig-vote_derniers_elections-overview
#| fig-cap: |
#|   Distribution of whether volunteers reported having voted or not during the last elections, averaged across time points and cohorts.
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

vote_plot <- combined |> 
  drop_na(vote_derniers_elections) |> 
  mutate(source = ifelse(source %in% c("q1", "q0"), "q0/q1", source)) |> mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  mutate(across(vote_derniers_elections, ~ factor(., levels = c("Pas sur les listes", "Non", "Oui")))) |> 
  group_by(vote_derniers_elections) |> 
  summarize(n_volunteers = n_distinct(id_jeune)) |> 
  ungroup() |> 
  mutate(share = n_volunteers/sum(n_volunteers)) 


ggplot(vote_plot, 
       aes(x = vote_derniers_elections, y = n_volunteers, fill = vote_derniers_elections)) +
  geom_col(position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]) +
  geom_text(aes(label = n_volunteers),
            position = position_dodge(width = 0.9),
            vjust = -0.3,
            size = 4) +
  labs(
    x = "Vote",
    y = "Nombre de volontaires",
    fill = "Vote lors des dernières élections"
  ) +
  guides(fill = FALSE) +
  plot_theme 
  
ggplot(vote_plot, 
       aes(x = vote_derniers_elections, y = share, fill = vote_derniers_elections)) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(aes(label = scales::percent(share, accuracy = 1)), 
            vjust = -0.5, size = 4) +
scale_fill_manual(values = wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Vote",
    y = "Nombre de volontaires",
    fill = "Vote lors des dernières élections"
  ) +
  guides(fill = FALSE) +
  plot_theme 
```


In the following sections, we investigate how volunteers have changed in answering this question, throughout their service civique. For these analyses, we only consider those volunteers who have answered the question at both time points, q1 and q2. Changes towards "Yes" or "No" can have happened either from the respective other option, or from the third answer option: "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales", which is abbreviated in the plots below as "Pas sur les listes". We look at the different cohorts separately (promo 2020-21, @fig-vote_derniers_elections-2020; promo 2021-22, @fig-vote_derniers_elections-2021; 2022-23, @fig-vote_derniers_elections-2022; 2023-24, @fig-vote_derniers_elections-2023). 

The following sections reveal considerable differences between the cohorts. For example, in 2020-21 and 2021-22, many volunteers reported changes in voting behavior, but in opposite directions: In 2020-21, 24% of volunteers changed towards "No", outweighing those who changed to "Yes" (6%) by far. Inversely, in 2021-22, 35% of volunteers changed towards "Yes", while only 8% changed towards "No".

It is not clear why we observe these differences. One explanation might be that volunteers care more about national elections than local ones (or vice verse). If that is the case, we would expect changes in voting behavior to be dependent on whether there were national or local elections happening during the year of a cohort's service civique.

### Promo 2020-21

@fig-vote_derniers_elections-2020 shows change in volunteers' reported voting behavior for the promo "2020-21".

```{r}
#| label: fig-vote_derniers_elections-2020
#| fig-cap: "Promo 2020-21. Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2


vote_data <- combined %>%
  filter(promo == "20-21") |> 
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

# run mcnemar test
#mcnemar.test(table(vote_data$vote_q1, vote_data$vote_q2))

# Prepare data for alluvial plot
vote_long <- vote_data |> 
  pivot_longer(cols = starts_with("vote_q"), names_to = "time", values_to = "vote") |> 
  mutate(time = recode(time, "vote_q1" = "q1", "vote_q2" = "q2"))


# Proportions
vote_long_prop <- vote_long |>
  group_by(time, vote) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(time) |>
  mutate(prop = n / sum(n),
         # make positioning variable
         position = sum(n) - n,
         ) |>
  ungroup()

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(vote_long, 
       aes(x = time, stratum = vote, alluvium = id_jeune, fill = vote, label = vote)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Vote"
  ) +
  plot_theme

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

```{r}
# for a plot that takes into account NAs

# vote_long <- combined %>%
#   filter(source %in% c("q1", "q2")) |>
#   select(id_jeune, source, vote_derniers_elections) %>%
#   rename(time = source, 
#          vote = vote_derniers_elections) |> 
#     mutate(vote = ifelse(is.na(vote), "Pas de réponse", vote)
#            )
# 
# # Create alluvial plot
# ggplot(vote_long, 
#        aes(x = time, stratum = vote, alluvium = id_jeune, fill = vote, label = vote)) +
#   geom_flow(stat = "alluvium",  alpha = 0.7) +
#   geom_stratum() +
#   scale_x_discrete(expand = c(0.1, 0.1)) +
#   labs(
#     title = "Changements dans le vote entre Q1 et Q2",
#     x = "Temps",
#     y = "Nombre d'individus",
#     fill = "Vote"
#   ) +
#   plot_theme
```

```{r}
# for a statistical test
# action_data <- combined %>%
#   drop_na(action_individuelle_societe) |>
#   select(id_jeune, source, action_individuelle_societe) |> 
#   rename(time = source) |> 
#   # select individuals with observations in both
#   group_by(id_jeune) |>
#   filter(all(c("q1", "q3") %in% time)) |>
#   ungroup() |> 
#   # make factors
#   mutate(
#     action_individuelle_societe = factor(
#       action_individuelle_societe,
#       levels = c("Pas vraiment", "Non", "Oui")  # ref = Pas vraiment
#     ),
#     time = factor(time, levels = c("q1", "q3"))
#   )
# 
# # Multinomial logistic mixed model
# mod <- brm(
#   action_individuelle_societe ~ time + (1 | id_jeune),
#   data = action_data,
#   family = categorical(),
#   chains = 4, cores = 4, seed = 1234
# )
# 
# exp(fixef(mod))  # For point estimates
# 
# library(ggeffects)
# 
# preds <- ggpredict(mod, terms = "time")
# plot(preds) +
#   labs(title = "Effet du programme sur les réponses",
#        x = "Temps", y = "Probabilité prédite")
```

### Promo 2021-22

@fig-vote_derniers_elections-2021 shows change in volunteers' reported voting behavior for the promo "2021-22".

```{r}
#| label: fig-vote_derniers_elections-2021
#| fig-cap: "Promo 2021-22. Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2


vote_data <- combined %>%
  filter(promo == "21-22") |> 
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

# run mcnemar test
#mcnemar.test(table(vote_data$vote_q1, vote_data$vote_q2))

# Prepare data for alluvial plot
vote_long <- vote_data |> 
  pivot_longer(cols = starts_with("vote_q"), names_to = "time", values_to = "vote") |> 
  mutate(time = recode(time, "vote_q1" = "q1", "vote_q2" = "q2"))


# Proportions
vote_long_prop <- vote_long |>
  group_by(time, vote) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(time) |>
  mutate(prop = n / sum(n),
         # make positioning variable
         position = sum(n) - n,
         ) |>
  ungroup()

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(vote_long, 
       aes(x = time, stratum = vote, alluvium = id_jeune, fill = vote, label = vote)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Vote"
  ) +
  plot_theme

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Promo 2022-23

@fig-vote_derniers_elections-2022 shows change in volunteers' reported voting behavior for the promo "2022-23".

```{r}
#| label: fig-vote_derniers_elections-2022
#| fig-cap: "Promo 2022-23. Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2

vote_data <- combined %>%
  filter(promo == "22-23") |> 
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

# run mcnemar test
#mcnemar.test(table(vote_data$vote_q1, vote_data$vote_q2))

# Prepare data for alluvial plot
vote_long <- vote_data |> 
  pivot_longer(cols = starts_with("vote_q"), names_to = "time", values_to = "vote") |> 
  mutate(time = recode(time, "vote_q1" = "q1", "vote_q2" = "q2"))


# Proportions
vote_long_prop <- vote_long |>
  group_by(time, vote) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(time) |>
  mutate(prop = n / sum(n),
         # make positioning variable
         position = sum(n) - n,
         ) |>
  ungroup()

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(vote_long, 
       aes(x = time, stratum = vote, alluvium = id_jeune, fill = vote, label = vote)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Vote"
  ) +
  plot_theme

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Promo 2023-24

@fig-vote_derniers_elections-2023 shows change in volunteers' reported voting behavior for the promo "2023-24".

```{r}
#| label: fig-vote_derniers_elections-2023
#| fig-cap: "Promo 2023-24. Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2

vote_data <- combined %>%
  filter(promo == "23-24") |> 
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

# run mcnemar test
#mcnemar.test(table(vote_data$vote_q1, vote_data$vote_q2))

# Prepare data for alluvial plot
vote_long <- vote_data |> 
  pivot_longer(cols = starts_with("vote_q"), names_to = "time", values_to = "vote") |> 
  mutate(time = recode(time, "vote_q1" = "q1", "vote_q2" = "q2"))


# Proportions
vote_long_prop <- vote_long |>
  group_by(time, vote) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(time) |>
  mutate(prop = n / sum(n),
         # make positioning variable
         position = sum(n) - n,
         ) |>
  ungroup()

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(vote_long, 
       aes(x = time, stratum = vote, alluvium = id_jeune, fill = vote, label = vote)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Vote"
  ) +
  plot_theme

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Which demographic factors are associated with change of voting intentions? 

To determine which demographic variables are associated with change of voting intentions, we ran statistical models. These models are explained in the Appendix (@sec-appendix-voting). Here, we only report the variables that were statistically significant in these models. 

```{r}
# remove variables that don't make sense, i.e. the motive and the actual duration (obviously, the longer the duration, the less likely to have a rupture)
reduced_demographic_variables <- setdiff(demographic_variables, c("motif_rupture"))

# add the two non-demographic variables asked in q1
factors_for_model <- c(reduced_demographic_variables, "programme_cle")

# make vote data
vote_data <- combined %>%
  drop_na(vote_derniers_elections) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections,  all_of(factors_for_model)) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui")))) |> 
  mutate(change_vote = case_when(
    vote_q1 == vote_q2 ~ NA_character_,
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  ), 
  change_vote = factor(change_vote, levels = c("Changé vers Non", "Changé vers Oui"))
  ) 

# Fit one model per demographic variable
model_results_voting_demographic <- factors_for_model  %>%
  map_dfr(function(var) {
    formula <- as.formula(glue::glue("change_vote ~ `{var}`"))
    
    model <- glm(formula, data = vote_data, family = "binomial")
    
    tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var, 
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })

# check
# model_results_voting_demographic |> 
#   filter(significant)
```

#### Sex

As shown in @fig-vote_derniers_elections-sex, more female volunteers changed towards "Yes" than towards "No". Among male volunteers, the opposite was true. 

```{r}
#| label: fig-vote_derniers_elections-sex
#| fig-cap: "Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2, according to their sex. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

vote_data <- combined %>%
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections, sexe) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, sexe) |> 
  summarize(n = n()) |>
  group_by(sexe) |> 
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~sexe) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# PERCENTAGES

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, sexe) |> 
  summarize(n = n()) |>
  group_by(sexe) |> 
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~sexe) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

#### Age

@fig-vote_derniers_elections-age compares changes in voting behaviour between age groups. Among the 18 to 20 year-olds, the share of volunteers who have changed towards "Yes" was highest. One interpretation could be, that the service civique might encourage this age group in particular to vote. However, an alternative explanation could be that this group could vote for the first time during their service civique, as the legal voting age in France is 18. 

```{r}
#| label: fig-vote_derniers_elections-age
#| fig-cap: "Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2, for the different age groups. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

vote_data <- combined %>%
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections, age_categorie) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, age_categorie) |> 
  summarize(n = n()) |>
  group_by(age_categorie) |> 
  mutate(prop = n / sum(n), 
         age_categorie = paste0("Age: ", age_categorie))

ggplot(vote_change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~age_categorie) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# PERCENTAGES

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, age_categorie) |> 
  summarize(n = n()) |>
  group_by(age_categorie) |> 
  mutate(prop = n / sum(n), 
         age_categorie = paste0("Age: ", age_categorie))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~age_categorie) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

On a side note, looking at the 16 to 17 age groups shows that some answers were erroneous. In France, the legal voting age is 18. Therefore, the 16 to 17 year-olds, at least at the beginning of their service civique (q0/q1), have never been on a voting list before. However, as shown in @fig-vote_derniers_elections-youngest there are 443 volunteers who answered "No" at q0/q1 (which technically is true, but not the most specific answer option), and 20 who answered "Yes" (probably not reading the question carefully or accidentally clicking on the wrong response). 

```{r}
#| label: fig-vote_derniers_elections-youngest
#| fig-cap: |
#|   Distribution of whether volunteers reported having voted or not during the last elections,
#|   at different time points, for the youngest age group (16–17 years old).
#|   All answers at q0/q1 that are not “Pas sur les listes” are not exactly correct.


combined |> 
  drop_na(vote_derniers_elections) |> 
  mutate(source = ifelse(source %in% c("q1", "q0"), "q0/q1", source)) |> mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  group_by(age_categorie, vote_derniers_elections, source) |> 
  summarize(n_volunteers = n_distinct(id_jeune), .groups = "drop") |> 
  filter(age_categorie == "16-17") |> 
  ggplot(aes(x = source, y = n_volunteers, fill = vote_derniers_elections)) +
  geom_col(position = position_dodge(width = 0.9)) +
  scale_fill_manual(values = wes_palette("Darjeeling1")[c(1, 3, 5)]) +
  geom_text(aes(label = n_volunteers),
            position = position_dodge(width = 0.9),
            vjust = -0.3,
            size = 3) +
  facet_wrap(~age_categorie) +
  labs(
    x = "Source",
    y = "Nombre de volontaires",
    fill = "Vote lors des dernières élections"
  ) +
  plot_theme +
  theme(legend.position = "top")
```

#### Zone of Residence

@fig-vote_derniers_elections-residence shows that volunteers from "QPV" and ZRR" residential areas had a greater tendency to change towards "No", compared to "Yes". By contrast, in particular in rural areas, but also slightly in urban ones, more volunteers changed towards "Yes". 

```{r}
#| label: fig-vote_derniers_elections-residence
#| fig-cap: "Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2, according to their zone of residence. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

vote_data <- combined %>%
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections, zone_residence) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, zone_residence) |> 
  summarize(n = n()) |>
  group_by(zone_residence) |> 
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~zone_residence) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# PERCENTAGES

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~zone_residence) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

#### Education

As shown in @fig-vote_derniers_elections-education, the share of volunteers who changed towards "Yes" was particularly high among those with a level of education corresponding to "Bac+2". In all other education segments, the share of those changing towards "No" was higher. 

```{r}
#| label: fig-vote_derniers_elections-education
#| fig-cap: "Change in volunteers reporting whether they voted or not during the last elections, between Q1 and Q2, according to their level of education. Note that this analysis considers only answers of volunteers who answered either yes or no at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

vote_data <- combined %>%
  drop_na(vote_derniers_elections) |>
  # filter(vote_derniers_elections %in% c("Oui", "Non")) |>
  mutate(vote_derniers_elections = case_when(
    vote_derniers_elections == "Vous n'aviez pas l'âge de voter ou vous n'étiez pas inscrit.e sur les listes électorales" ~ "Pas sur les listes", 
    TRUE ~ vote_derniers_elections
  )
  ) |> 
  select(id_jeune, source, vote_derniers_elections, niveau_etudes) %>%
  pivot_wider(names_from = source, values_from = vote_derniers_elections, names_prefix = "vote_") %>%
  filter(!is.na(vote_q1) & !is.na(vote_q2)) %>%
  mutate(across(starts_with("vote_"), ~ factor(., levels = c("Pas sur les listes", "Non", "Oui"))))

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, niveau_etudes) |> 
  summarize(n = n()) |>
  group_by(niveau_etudes) |> 
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~niveau_etudes) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

# PERCENTAGES

vote_change_summary <- vote_data |>
  mutate(change = case_when(
    vote_q1 == vote_q2 ~ "Pas de changement",
    vote_q1 == "Oui" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Non" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    vote_q1 == "Pas sur les listes" & vote_q2 == "Non" ~ "Changé vers Non",
    vote_q1 == "Pas sur les listes" & vote_q2 == "Oui" ~ "Changé vers Oui", 
    TRUE ~ NA_character_
  )) |>
  drop_na(change) |> 
  group_by(change, niveau_etudes) |> 
  summarize(n = n()) |>
  group_by(niveau_etudes) |> 
  mutate(prop = n / sum(n))

ggplot(vote_change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 0.1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    #"Pas de changement" = "#E1AF00",
    "Changé vers Non" = "#F21A00",
    "Changé vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    #title = "Changements individuels entre Q1 et Q2",
    x = NULL,
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~niveau_etudes) +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

## Individual action for society {#sec-individual-action}

In this section, we look at how volunteers have changed regarding their perception on whether their individual action can contribute to changing society. This was measured with the question: "En général, pensez-vous que votre action individuelle peut contribuer à changer la société ?", which has been asked at time points q1 and q3. We look at the different cohorts separately (promo 2020-21, @fig-action_individuelle_societe-2020; promo 2021-22, @fig-vote_derniers_elections-2021; 2022-23, @fig-vote_derniers_elections-2022; 2023-24, @fig-vote_derniers_elections-2023).

Descriptively, there is no clear positive or negative trend. On average, across all cohorts, slightly more people are changing towards "No" or "Not really", compared to "Yes".

### Promo 2020-21

@fig-action_individuelle_societe-2020 shows change in volunteers' perceptions of whether their individual action contributes to changing society, for the promo "2020-21".

```{r}
#| label: fig-action_individuelle_societe-2020
#| fig-cap: "Promo 2020-21. Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3. Note that this analysis considers only answers of volunteers who answered at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2

# make wide format data to track change
action_data <- combined %>%
  filter(promo == "20-21") |> 
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))


# Prepare data for alluvial plot
plot_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) |> 
  rename(time = source) |> 
  # select individuals with observations in both
  group_by(id_jeune) |>
  filter(all(c("q1", "q3") %in% time)) |>
  ungroup() |> 
  # make a factor
    mutate(action_individuelle_societe = factor(
    action_individuelle_societe,
    levels = c("Pas vraiment", "Non", "Oui")
  ))

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(plot_data, 
       aes(x = time, stratum = action_individuelle_societe, alluvium = id_jeune, fill = action_individuelle_societe, label = action_individuelle_societe)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Contribution individuelle"
  ) +
  plot_theme 

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Promo 2021-22

@fig-action_individuelle_societe-2021 shows change in volunteers' perceptions of whether their individual action contributes to changing society, for the promo "2021-22".

```{r}
#| label: fig-action_individuelle_societe-2021
#| fig-cap: "Promo 2021-22. Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3. Note that this analysis considers only answers of volunteers who answered at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2

# make wide format data to track change
action_data <- combined %>%
  filter(promo == "21-22") |> 
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))


# Prepare data for alluvial plot
plot_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) |> 
  rename(time = source) |> 
  # select individuals with observations in both
  group_by(id_jeune) |>
  filter(all(c("q1", "q3") %in% time)) |>
  ungroup() |> 
  # make a factor
    mutate(action_individuelle_societe = factor(
    action_individuelle_societe,
    levels = c("Pas vraiment", "Non", "Oui")
  ))

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(plot_data, 
       aes(x = time, stratum = action_individuelle_societe, alluvium = id_jeune, fill = action_individuelle_societe, label = action_individuelle_societe)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Contribution individuelle"
  ) +
  plot_theme 

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Promo 2022-23

@fig-action_individuelle_societe-2022 shows change in volunteers' perceptions of whether their individual action contributes to changing society, for the promo "2022-23".

```{r}
#| label: fig-action_individuelle_societe-2022
#| fig-cap: "Promo 2022-23. Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3. Note that this analysis considers only answers of volunteers who answered at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2

# make wide format data to track change
action_data <- combined %>%
  filter(promo == "22-23") |> 
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))


# Prepare data for alluvial plot
plot_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) |> 
  rename(time = source) |> 
  # select individuals with observations in both
  group_by(id_jeune) |>
  filter(all(c("q1", "q3") %in% time)) |>
  ungroup() |> 
  # make a factor
    mutate(action_individuelle_societe = factor(
    action_individuelle_societe,
    levels = c("Pas vraiment", "Non", "Oui")
  ))

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(plot_data, 
       aes(x = time, stratum = action_individuelle_societe, alluvium = id_jeune, fill = action_individuelle_societe, label = action_individuelle_societe)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Contribution individuelle"
  ) +
  plot_theme 

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Promo 2023-24

@fig-action_individuelle_societe-2023 shows change in volunteers' perceptions of whether their individual action contributes to changing society, for the promo "2023-24".

```{r}
#| label: fig-action_individuelle_societe-2023
#| fig-cap: "Promo 2023-24. Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3. Note that this analysis considers only answers of volunteers who answered at *both* time points."
#| fig-subcap:
#|   - "Alluvial plot"
#|   - "Percentages" 
#| layout-ncol: 2

# make wide format data to track change
action_data <- combined %>%
  filter(promo == "23-24") |> 
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))


# Prepare data for alluvial plot
plot_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe) |> 
  rename(time = source) |> 
  # select individuals with observations in both
  group_by(id_jeune) |>
  filter(all(c("q1", "q3") %in% time)) |>
  ungroup() |> 
  # make a factor
    mutate(action_individuelle_societe = factor(
    action_individuelle_societe,
    levels = c("Pas vraiment", "Non", "Oui")
  ))

pal <- wes_palette("Zissou1", 5, type = "discrete")[c(3, 5, 1)]

# Create alluvial plot
ggplot(plot_data, 
       aes(x = time, stratum = action_individuelle_societe, alluvium = id_jeune, fill = action_individuelle_societe, label = action_individuelle_societe)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.7) +
  geom_stratum() +
  scale_fill_manual(values = pal) + 
  scale_x_discrete(expand = c(0.1, 0.1)) +
  labs(
    x = "Temps",
    y = "Nombre d'individus",
    fill = "Contribution individuelle"
  ) +
  plot_theme 

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  count(change) |>
  mutate(prop = n / sum(n))

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.1))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme
```

### Which demographic factors are associated with change in perceptions of one's individual contribution to society? 

Whether volunteers perceived their individual action can contribute to changing society varies along different demographic variables. These models are explained in the Appendix (@sec-appendix-individual-action). Here, we only report the variables that were statistically significant in these models. 

```{r}
# remove variables that don't make sense, i.e. the motive and the actual duration (obviously, the longer the duration, the less likely to have a rupture)
reduced_demographic_variables <- setdiff(demographic_variables, c("motif_rupture"))

# add the two non-demographic variables asked in q1
factors_for_model <- c(reduced_demographic_variables, "programme_cle")

# make action data
action_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe, all_of(factors_for_model)) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3)) |>
  # make a binary version of change
  mutate(change_action = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ NA_character_,
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non/Pas vraiment",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ NA_character_,
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Non/Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ NA_character_
  ), 
  change_action = factor(change_action, levels = c("vers Non/Pas vraiment", "vers Oui")
  )
  )


# Fit one model per demographic variable
model_results_action <- factors_for_model  %>%
  map_dfr(function(var) {
    formula <- as.formula(glue::glue("change_action ~ `{var}`"))
    
    model <- glm(formula, data = action_data, family = "binomial")
    
    tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE)
             )
  })

# check
# model_results_action |>
#   filter(significant)
```

#### Zone of Residence

@fig-action_individuelle_societe-residence shows that in particular volunteers from rural areas had a greater tendency to change towards "Yes", compared to changes towards "No" or "Not really". 

```{r}
#| label: fig-action_individuelle_societe-residence
#| fig-cap: "Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3, according to their zone of residence. Note that this analysis considers only answers of volunteers who provided an answer at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# make wide format data to track change
action_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe, zone_residence) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  group_by(change, zone_residence) |> 
  summarize(n = n()) |>
  group_by(zone_residence) |>
  mutate(prop = n / sum(n))

ggplot(change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~zone_residence) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))

# PERCENTAGES

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.3))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~zone_residence) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

#### Education

@fig-action_individuelle_societe-education shows that volunteers with a "Bac+3" had a greater tendency to change towards "Yes", compared to changing towards "No" or "Not really", than the other groups. 

```{r}
#| label: fig-action_individuelle_societe-education
#| fig-cap: "Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3, according to their level of education. Note that this analysis considers only answers of volunteers who provided an answer at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# make wide format data to track change
action_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe, niveau_etudes) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  group_by(change, niveau_etudes) |> 
  summarize(n = n()) |>
  group_by(niveau_etudes) |>
  mutate(prop = n / sum(n))

ggplot(change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~niveau_etudes) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))

# PERCENTAGES

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.3))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~niveau_etudes) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

#### Handicap (administrative)

@fig-action_individuelle_societe-handicap shows that volunteers with a handicap (administrative) had a greater tendency to change towards "No" or "Not really", compared to changing towards "Yes". 

```{r}
#| label: fig-action_individuelle_societe-handicap
#| fig-cap: "Change in volunteers reporting whether they think their individual action can contribute to changing society, between Q1 and Q3, according to whether they had a handicap (administrative) or not. Note that this analysis considers only answers of volunteers who provided an answer at *both* time points."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# make wide format data to track change
action_data <- combined %>%
  drop_na(action_individuelle_societe) |>
  select(id_jeune, source, action_individuelle_societe, handicap_administratif) %>%
  pivot_wider(names_from = source, values_from = action_individuelle_societe, names_prefix = "action_individuelle_") %>%
  filter(!is.na(action_individuelle_q1) & !is.na(action_individuelle_q3))

# make plot for change

change_summary <- action_data |>
  mutate(change = case_when(
    action_individuelle_q1 == action_individuelle_q3 ~ "Pas de changement",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Non" ~ "vers Non",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Pas vraiment" & action_individuelle_q3 == "Oui" ~ "vers Oui",
    action_individuelle_q1 == "Oui" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment",
    action_individuelle_q1 == "Non" & action_individuelle_q3 == "Pas vraiment" ~ "vers Pas vraiment"
  )) |>
  group_by(change, handicap_administratif) |> 
  summarize(n = n()) |>
  group_by(handicap_administratif) |>
  mutate(prop = n / sum(n)) |> 
  mutate(handicap_administratif = paste0("Handicap: ", handicap_administratif))

ggplot(change_summary, aes(x = change, y = n, fill = change)) +
  geom_col() +
  geom_text(aes(label = n), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.2))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~handicap_administratif) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))

# PERCENTAGES

ggplot(change_summary, aes(x = change, y = prop, fill = change)) +
  geom_col() +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            vjust = -0.5, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.3))) +
  scale_fill_manual(values = c(
    "vers Pas vraiment" = "#E1AF00",
    "vers Non" = "#F21A00",
    "vers Oui" = "#3B9AB2"
  )) +
  guides(fill = FALSE) +
  labs(
    x = "Changé",
    y = "Proportion",
    fill = NULL
  ) +
  plot_theme +
  facet_wrap(~handicap_administratif) +
  theme(axis.text.x = element_text(angle = 40, hjust = 1))
```

# Rupture {#sec-rupture}

```{r}
share_rupture <- combined |> 
  group_by(rupture) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = paste0(round(count / sum(count) * 100, 1), "%")) %>%
  split(.$rupture)
```

Not all volunteers work until the end of their contract. In fact, `r share_rupture$rupture$share` of volunteers have a "rupture", i.e. terminate the contract early (@fig-rupture-overview). There are various motives for ending one's contract early (@fig-rupture-motive). Not all of them are necessarily bad, e.g. "Embauche en CDD d'au moins 6 mois ou CDI", and some are outside of the influence of the volunteers, e.g. "Fin de validité du Titre de Séjour". 

```{r}
#| label: fig-rupture-overview
#| fig-cap: "Number of volunteers with a rupture (for various possible reasons, including positive ones, such as obtaining a work contract), pooled across all cohorts."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

combined |> 
  group_by(rupture) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = rupture, y = count, fill = rupture)) +
  geom_col() +
  scale_fill_viridis_d(option = "plasma") +
  geom_text(aes(label = count), vjust = -0.5, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Rupture",
    y = "Number of volunteers"
  ) +
  plot_theme +
  guides(fill = FALSE)

combined |> 
  group_by(rupture) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = rupture, y = share, fill = rupture)) +
  geom_col() +
  scale_fill_viridis_d(option = "plasma") +
  geom_text(aes(label = paste0(share*100, "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Rupture",
    y = "Share of volunteers"
  ) +
  plot_theme +
  guides(fill = FALSE)
```


```{r}
#| label: fig-rupture-motive
#| fig-cap: "Prevalence of different rupture motives, pooled across all cohorts."

combined |> 
  group_by(motif_rupture, rupture_valence) |> 
  drop_na(motif_rupture) |> 
  summarise(n = n_distinct(id_jeune), .groups = "drop") |> 
  mutate(motif_rupture = fct_reorder(motif_rupture, n, .fun = sum)) |> 
  ggplot(aes(x = motif_rupture, y = n, fill = rupture_valence)) +
  geom_col(position = "dodge") +
  geom_text(aes(label = n), 
            position = position_dodge(width = 0.9),
            vjust = 0.3, hjust = -0.1, size = 3) +
  scale_fill_viridis_d(option = "plasma", end = 0.85) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = NULL,
    y = "Number of volunteers",
    fill = "Valence de la rupture"
  ) +
  labs(fill = NULL) +
  coord_flip() + 
  theme(legend.position = "top")
  
```


## Change in rupture

@fig-rupture illustrates change regarding how many volunteers have ended their contract early (rupture) by comparing the different promos. There are no clear trends with slightly more than 20% of volunteers having a rupture.

```{r}
#| label: fig-rupture
#| fig-cap: "Number of volunteers with a rupture (for various possible reasons, including positive ones, such as obtaining a work contract)."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

combined |> 
  group_by(promo, rupture) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = rupture, y = count, fill = rupture)) +
  geom_col() +
  scale_fill_viridis_d(option = "plasma") +
  geom_text(aes(label = count), vjust = -0.5, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Rupture",
    y = "Number of volunteers"
  ) +
  plot_theme +
  facet_wrap(~promo) +
  guides(fill = FALSE)

combined |> 
  group_by(promo, rupture) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = rupture, y = share, fill = rupture)) +
  geom_col() +
  scale_fill_viridis_d(option = "plasma") +
  geom_text(aes(label = paste0(share*100, "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Rupture",
    y = "Share of volunteers"
  ) +
  plot_theme +
  facet_wrap(~promo) +
  guides(fill = FALSE)
  
```

## What predicts whether volunteers end their contract early (rupture)?

Statistical models on rupture can be found in the Appendix (@sec-appendix-rupture). Here, we only report the variables that were statistically significant in these models. Not all ruptures happen for negative reasons (@fig-rupture-motive). For analyses in this section, we focus on negative reasons: We distinguish between volunteers who ended their contract early for apparently *negative* reasons, and all other volunteers. 

```{r}
# remove variables that don't make sense, e.g., the motive and the actual duration (obviously, the longer the duration, the less likely to have a rupture)
reduced_demographic_variables <- setdiff(demographic_variables, c("rupture", "motif_rupture", "Duree_reelle_sc_mois"))

# add the two non-demographic variables asked in q1
factors_for_rupture <- c("perception_avenir_num", "action_individuelle_societe")

# make model data - we only want to have one rupture value per volunteer
rupture_data <- combined |>
  arrange(id_jeune, is.na(motif_rupture)) |>  # NA last
  group_by(id_jeune) |>
  slice(1) |>  # Keep first (i.e. non-NA if it exists)
  ungroup()

# Fit one model per demographic variable
model_results_rupture_other_variables <- factors_for_rupture %>%
  map_dfr(function(var) {
    df <- rupture_data %>%
      select(rupture_negative, !!sym(var)) %>%
      drop_na()
    
    # Only run if predictor has 2+ unique levels
    if (n_distinct(df[[var]]) < 2) {
      return(NULL)
    }
    
    formula <- as.formula(glue::glue("rupture_negative ~ `{var}`"))
    
    model <- glm(formula, data = df, family = "binomial")
    
    tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })

# check
# model_results_other_variables

# Fit one model per demographic variable
model_results_rupture_demographic_variables <- reduced_demographic_variables %>%
  map_dfr(function(var) {
    df <- rupture_data %>%
      select(rupture_negative, !!sym(var)) %>%
      drop_na()
    
    # Only run if predictor has 2+ unique levels
    if (n_distinct(df[[var]]) < 2) {
      return(NULL)
    }
    
    formula <- as.formula(glue::glue("rupture_negative ~ `{var}`"))
    
    model <- glm(formula, data = df, family = "binomial")
    
    tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })



# check
# model_results_rupture_demographic_variables |> 
#   filter(significant)
# 
# model_results_rupture_other_variables |> 
#   filter(significant)

```

### Demographic factors

As shown in @fig-rupture-descriptive, older volunteers tend to have less ruptures than younger ones; less educated volunteers (in particular "infra bac") tend to have more ruptures than those with higher education; refugees tend to have less ruptures than non-refugees; male volunteers tend to have slightly more ruptures than female ones; volunteers in the "Relai" program tend to have slightly less ruptures than volunteers in the "Coeur" program; volunteers from "QPV" residential areas tend to have considerably higher rupture rates than volunteers from other residential areas. 

```{r}
#| label: fig-rupture-descriptive 
#| fig-cap: Percentages of rupture for (allegedly) negative reasons for different groups, for different variables.

plot_variables <- c("age_categorie", "type_volontaire", "niveau_etudes", "sexe", "refugie", "zone_residence")

plot_data <- combined %>%
  mutate(across(all_of(plot_variables), as.character)) %>%
  pivot_longer(cols = all_of(plot_variables), names_to = "variable", values_to = "level") %>%
  drop_na(level) |> 
  group_by(variable, level, rupture_negative) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(variable, level) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# add correct levels for order
level_orders <- lapply(combined[plot_variables], levels)

plot_data <- plot_data %>%
  group_by(variable) %>%
  mutate(level = factor(level, levels = level_orders[[unique(variable)]])) %>%
  ungroup()

ggplot(plot_data, aes(x = level, y = pct, fill = factor(rupture_negative))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_viridis_d(option = "rocket", direction = -1) +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, y = "Share",
    fill = "Rupture",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct) * 1.2)) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1), 
        legend.position = "top")
```

### Other factors

For non-demographic variables, investigating their relationship with rupture is not possible--simply because, by definition, for questions that have been only asked at "q2" and "q3", volunteers who had ended their contract early were not available anymore (see @tbl-rupture-candidates). Only for the two variables that have been asked at "q1": 

1. `perception_avenir` (Concernant votre avenir, êtes-vous...? 	Très confiant(e); Assez confiant(e); Peu confiant(e); Pas du tout confiant(e))
2. `action_individuelle_societe` (En général, pensez-vous que votre action individuelle peut contribuer à changer la société ? 	Oui; Pas vraiment; Non)

As shown in @fig-rupture-other, perhaps unsurprisingly, those volunteers who tend to think that their individual action can contribute to changing society tend to disrupt their contract less. Interestingly, volunteers who are very confident regarding their future tend to distrupt their contract more. 

```{r}
#| label: fig-rupture-other
#| fig-cap: Percentages of rupture for (allegedly) negative reasons, as a function of (i) the volunteers' confidence in their future and (ii) their perception that their individual action can contribute to changing society.

plot_variables <- c("perception_avenir", "action_individuelle_societe")

plot_data <- combined %>%
  mutate(across(all_of(plot_variables), as.character)) %>%
  pivot_longer(cols = all_of(plot_variables), names_to = "variable", values_to = "level") %>%
  drop_na(level) |> 
  mutate(level = factor(level)) %>%
  group_by(variable, level, rupture_negative) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(variable, level) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# add correct levels for order
level_orders <- lapply(combined[plot_variables], levels)

plot_data <- plot_data %>%
  group_by(variable) %>%
  mutate(level = factor(level, levels = level_orders[[unique(variable)]])) %>%
  ungroup()

ggplot(plot_data, aes(x = level, y = pct, fill = factor(rupture_negative))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_viridis_d(option = "rocket", direction = -1) +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, y = "Share",
    fill = "Rupture",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct) * 1.2)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1), 
        legend.position = "top")
```

# Satisfaction {#sec-satisfaction}

In this section, we look at satisfaction ("D'une manière générale, diriez-vous que votre Service Civique s'est déroulé de façon..." with levels 1, "pas du tout satisfaisainte", to 4, "très satisfaisante")^[In all analyses we treat this as a continuous variable]. As shown in @fig-satisfaction-overview, taking all cohorts together, the majority of volunteers thinks their experience is "très satisfaisant". 

```{r}
#| label: fig-satisfaction-overview
#| fig-cap: "Répartition des niveaux de satisfaction" 
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

combined |> 
  drop_na(satisfaction) |> 
  group_by(satisfaction) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = satisfaction, y = count)) +
  geom_col() +
  geom_text(aes(label = count), vjust = -0.5, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Satisfaction",
    y = "Number of volunteers"
  ) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))

color <- wes_palette("Moonrise3")[2]

combined |> 
  drop_na(satisfaction) |> 
  group_by(satisfaction) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = satisfaction, y = share)) +
  geom_col(fill = color) +
  geom_text(aes(label = paste0(share*100, "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Satisfaction",
    y = "Share of volunteers"
  ) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))
```

## Change in Satisfaction

@fig-satisfaction shows how satisfaction with the service civique has changed over time, between the different cohorts. The share of volunteers who are "very satisfied" has decreased over time, but the shares of those who are "not" or "little" satisfied have remained broadly stable. 

```{r}
#| label: fig-satisfaction
#| fig-cap: "Satisfaction between cohorts." 
#| 
combined <- combined |> 
  # make satisfaction a factor, also make a numeric version
  mutate(
    satisfaction = factor(
      satisfaction,
      levels = c(
        "Pas du tout satisfaisante",
        "Peu satisfaisante",
        "Assez satisfaisante",
        "Très satisfaisante"
      ),
      ordered = TRUE
    ),
    satisfaction_num = as.numeric(satisfaction)
  )

color <- wes_palette("Moonrise3")[2]

combined |> 
  drop_na(satisfaction) |> 
  count(promo, satisfaction) |> 
  group_by(promo) |> 
  mutate(prop = n / sum(n)) |> 
  ggplot(aes(x = satisfaction, y = prop)) +
  geom_col(position = "dodge", fill = color) +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5, size = 3) +
    scale_y_continuous(limits = c(0, NA), 
                       expand = expansion(mult = c(0, 0.1)),
                       labels = scales::percent_format(accuracy = 1)
                       ) +
  labs(
    x = "Satisfaction",
    y = "Proportion",
    fill = "Promo",
    title = "Répartition des niveaux de satisfaction par promo"
  ) +
  guides(fill = FALSE) + 
  facet_wrap(~promo) +  # Facet by promo
  plot_theme +
  theme(
    axis.text.x = element_text(angle = 20, hjust = 1)  # Rotate x-axis labels
  )

```

## What predicts whether volunteers are more satisfied ?

Statistical models on satisfaction can be found in the Appendix (@sec-appendix-satisfaction). Here, we only report the variables that were statistically significant in these models. 

```{r}
# non-demographic variables 
factors_for_satisfaction <- c(
  "perception_avenir_num",
  "action_individuelle_societe",
  "comparaison_utilite_autres",
  "fierte_num",
  "confiance_en_soi_num",
  "confiance_avenir_personnel_num",
  "impact_situation_actuelle"
)


# make model data - we only want to have one satisfaction value per volunteer
satisfaction_data <- combined |>
  arrange(id_jeune, is.na(satisfaction)) |>  # NA last
  group_by(id_jeune) |>
  slice(1) |>  # Keep first (i.e. non-NA if it exists)
  ungroup()

# Fit one model per other variable
model_results_satisfaction_other_variables <- factors_for_satisfaction %>%
  map_dfr(function(var) {
    df <- satisfaction_data %>%
      select(satisfaction_num, !!sym(var)) %>%
      drop_na()

    # Only run if predictor has 2+ unique levels
    if (n_distinct(df[[var]]) < 2) {
      return(NULL)
    }

    formula <- as.formula(glue::glue("satisfaction_num ~ `{var}`"))
    model <- lm(formula, data = df)

    tidy(model, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })

# Fit one model per demographic variable
model_results_satisfaction_demographic_variables <- demographic_variables %>%
  map_dfr(function(var) {
    df <- satisfaction_data %>%
      select(satisfaction_num, !!sym(var)) %>%
      drop_na()

    # Only run if predictor has 2+ unique levels
    if (n_distinct(df[[var]]) < 2) {
      return(NULL)
    }

    formula <- as.formula(glue::glue("satisfaction_num ~ `{var}`"))
    model <- lm(formula, data = df)

    tidy(model, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })


# check
# model_results_satisfaction_demographic_variables |> 
#   filter(significant)
# 
# model_results_satisfaction_other_variables |> 
#   filter(significant)

```

### Demographic factors

As shown in @fig-satisfaction-descriptive, older volunteers tend to be more satisfied than younger ones; more educated volunteers tend to be more satisfied than less educated; refugees tend to have less ruptures than non-refugees; volunteers in the "Relai" program tend to be more satisfied than volunteers in the "Coeur" program; probably unsurprisingly, volunteers who have a rupture tend to be less satisfied than volunteers who don't. 

```{r}
#| label: fig-satisfaction-descriptive 
#| fig-cap: Share of volunteers responding to the different satisfaction levels, grouped by different variables.

plot_variables <- c("rupture", "age_categorie", "type_volontaire", "niveau_etudes")

plot_data <- combined %>%
  mutate(across(all_of(plot_variables), as.character)) %>%
  pivot_longer(cols = all_of(plot_variables), names_to = "variable", values_to = "level") %>%
  drop_na(level) |> 
  drop_na(satisfaction) |> 
  group_by(variable, level, satisfaction) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(variable, level) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# add correct levels for order
level_orders <- lapply(combined[plot_variables], levels)

plot_data <- plot_data %>%
  group_by(variable) %>%
  mutate(level = factor(level, levels = level_orders[[unique(variable)]])) %>%
  ungroup()

# set colours
ggplot(plot_data, aes(x = level, y = pct, fill = factor(satisfaction))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Blues") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, y = "Share",
    fill = "Satisfaction",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct))) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1), 
        legend.position = "top")
```

### Other factors

The following variables are statistically associated with satisfaction (@fig-satisfaction-other): 

- `comparaison_utilite_autres` ["Diriez-vous que votre Service Civique a été...? 	Plus utile à vous qu'aux autres; Autant utile à vous qu'aux autres; Moins utile à vous qu'aux autres"]. Probably unsurprisingly, volunteers who believe the service civique was less useful for them than for others tend to be less satisfied than those who think it was more usefult to them 
- `fierte` ["Vous sentez-vous fier de ce que vous avez accompli pendant votre Service Civique ? 	Oui, tout à fait; Oui, un peu; Non, pas vraiment; Non, pas du tout"]. Volunteers who are more proud with what they have achieved during their service civique also tend to be more satisfied. 
- `confiance_en_soi` ["Pensez-vous avoir gagné en confiance en vous durant votre Service Civique ? 	Oui, beaucoup; Oui, assez; Non, pas vraiment; Non, pas du tout; Je ne sais pas"] Volunteers who have gained confidence also tend to be more satisfied.
- `confiance_avenir_personnel` ["Concernant votre avenir personnel, êtes-vous...? 	Très confiant.e; Assez confiant.e; Peu confiant.e; Pas du tout confiant.e"] Volunteers who are more confident about their future also tend to be more satisfied. 

```{r}
#| label: fig-satisfaction-other
#| fig-cap: Share of volunteers responding to the different satisfaction levels, grouped by different variables.

plot_variables <- c(
  "comparaison_utilite_autres",
  "fierte",
  "confiance_en_soi",
  "confiance_avenir_personnel"
)

plot_data <- combined %>%
  mutate(across(all_of(plot_variables), as.character)) %>%
  pivot_longer(cols = all_of(plot_variables), names_to = "variable", values_to = "level") %>%
  drop_na(level) |> 
  drop_na(satisfaction) |> 
  group_by(variable, level, satisfaction) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(variable, level) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# add correct levels for order
level_orders <- lapply(combined[plot_variables], levels)

plot_data <- plot_data %>%
  group_by(variable) %>%
  mutate(level = factor(level, levels = level_orders[[unique(variable)]])) %>%
  ungroup()

# set colours
ggplot(plot_data, aes(x = level, y = pct, fill = factor(satisfaction))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Blues") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, y = "Share",
    fill = "Satisfaction",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct))) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1), 
        legend.position = "top")
```

## Differences between key programs

@fig-satisfaction-programs shows descriptive differences in satisfaction between different key programs. 

```{r}
#| label: fig-satisfaction-programs
#| fig-cap: "Share of volunteers responding to the different satisfaction levels, grouped by which program they were in. Note that we focus only on key programs here. All other programs appear under 'NA' in the plot."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

plot_data <- combined %>%
  drop_na(satisfaction) |> 
  group_by(programme_cle, satisfaction) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(programme_cle) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# absolute numbers
ggplot(plot_data, aes(x = satisfaction, y = n, fill = factor(satisfaction))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Blues") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, 
    y = "Number of volunteers",
    fill = "Satisfaction",
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.2))) +
  plot_theme +
  theme(axis.text.x = element_blank(), 
        legend.position = "top") +
  facet_wrap(~programme_cle)


# percentages
ggplot(plot_data, aes(x = satisfaction, y = pct, fill = factor(satisfaction))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Blues") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, 
    y = "Share of volunteers",
    fill = "Satisfaction",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct))) +
  plot_theme +
  theme(axis.text.x = element_blank(), 
        legend.position = "top") +
  facet_wrap(~programme_cle)

```

# Confidence in one's future

In this section, we look at confidence in one's future (`perception_avenir`: "Concernant votre avenir, êtes-vous...?" with levels 1, "Pas du tout confiant.e", to 4, "Très confiant.e")^[In all analyses we treat this as a continuous variable]. As shown in @fig-perception-avenir-overview, taking all cohorts together, the majority of volunteers are "assez confiant.e". 

```{r}
#| label: fig-perception-avenir-overview
#| fig-cap: "Confidence in one's future, pooled across all cohorts." 
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

combined |> 
  drop_na(perception_avenir) |> 
  group_by(perception_avenir) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = perception_avenir, y = count)) +
  geom_col() +
  geom_text(aes(label = count), vjust = -0.5, size = 3) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Confidence in one's future",
    y = "Number of volunteers"
  ) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1)) 

color <- wes_palette("Cavalcanti1")[4]

combined |> 
  drop_na(perception_avenir) |> 
  group_by(perception_avenir) |> 
  summarise(
    count = n_distinct(id_jeune)
  ) |> 
  mutate(share = round(count / sum(count), 3)
         ) |> 
  ggplot(aes(x = perception_avenir, y = share)) +
  geom_col(fill = color) +
  geom_text(aes(label = paste0(share*100, "%")), vjust = -0.5, size = 3) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Confidence in one's future",
    y = "Share of volunteers"
  ) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1))


```

## Change in confidence in one's future

@fig-perception_avenir shows how confidence in one's future has changed over time, between the different cohorts. The results are very stable, with a majority of volunteers being "assez confiant.e" in all cohorts.

```{r}
#| label: fig-perception_avenir
#| fig-cap: "Confidence in one's future, between cohorts." 

color <- wes_palette("Cavalcanti1")[4]

combined |> 
  drop_na(perception_avenir) |> 
  group_by(perception_avenir, promo) |> 
  summarise(n = n_distinct(id_jeune)) |> 
  group_by(promo) |> 
  mutate(prop = n / sum(n)) |> 
  ggplot(aes(x = perception_avenir, y = prop)) +
  geom_col(position = "dodge", fill = color) +
  geom_text(aes(label = scales::percent(prop, accuracy = 1)), 
            position = position_dodge(width = 0.8), 
            vjust = -0.5, size = 3) +
    scale_y_continuous(limits = c(0, NA), 
                       expand = expansion(mult = c(0, 0.1)),
                       labels = scales::percent_format(accuracy = 1)
                       ) +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "C", end = 0.85) +
  labs(
    x = "Perceptions of one's future",
    y = "Proportion"
  ) +
  plot_theme  + 
  facet_wrap(~promo) +  # Facet by promo
  plot_theme +
  theme(
    axis.text.x = element_text(angle = 15, hjust = 1)  # Rotate x-axis labels
  )

```

## What predicts whether volunteers are more confident in their future ? {#sec-perception-avenir}

Statistical models on volunteers' confidence in their future can be found in the Appendix (@sec-appendix-perception-avenir). Here, we only report the variables that were statistically significant in these models. 

```{r}
# non-demographic variables 
factors_for_perception_avenir <- c(
  "satisfaction_num",
  "action_individuelle_societe",
  "comparaison_utilite_autres",
  "fierte_num",
  "confiance_en_soi_num",
  "confiance_avenir_personnel_num",
  "impact_situation_actuelle"
)


# make model data - we only want to have one satisfaction value per volunteer
perception_avenir_data <- combined |>
  arrange(id_jeune, is.na(perception_avenir)) |>  # NA last
  group_by(id_jeune) |>
  slice(1) |>  # Keep first (i.e. non-NA if it exists)
  ungroup()

# Fit one model per other variable
model_results_perception_other_variables <- factors_for_perception_avenir %>%
  map_dfr(function(var) {
    df <- perception_avenir_data %>%
      select(perception_avenir_num, !!sym(var)) %>%
      drop_na()

    # Only run if predictor has 2+ unique levels
    if (n_distinct(df[[var]]) < 2) {
      return(NULL)
    }

    formula <- as.formula(glue::glue("perception_avenir_num ~ `{var}`"))
    model <- lm(formula, data = df)

    tidy(model, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })

# Fit one model per demographic variable
model_results_perception_demographic_variables <- demographic_variables %>%
  map_dfr(function(var) {
    df <- perception_avenir_data %>%
      select(perception_avenir_num, !!sym(var)) %>%
      drop_na()

    # Only run if predictor has 2+ unique levels
    if (n_distinct(df[[var]]) < 2) {
      return(NULL)
    }

    formula <- as.formula(glue::glue("perception_avenir_num ~ `{var}`"))
    model <- lm(formula, data = df)

    tidy(model, conf.int = TRUE) %>%
      filter(term != "(Intercept)") %>%
      mutate(variable = var,
             significant = ifelse(p.value <= 0.05, TRUE, FALSE))
  })


# check
# model_results_perception_demographic_variables |>
#   filter(significant)
# 
# model_results_perception_other_variables |>
#   filter(significant)
```

### Demographic factors

As shown in @fig-perception-avenir-descriptive, volunteers in the "Relai" program tend to be more confident about their future than volunteers in the "Coeur" program; male volunteers tend to be more confident than female ones; refugees tend to be more confident than non-refugees; compared to volunteers from urban residential zones, volunteers from "QPV" zones tend to be slightly more confident, while volunteers from "ZRR" zones tend to be slightly less confident; the more educated volunteers, the less confident they tend to be. For age, the pattern appears non-linear: volunteers from the age group "16-17" tend to be more confident than volunteers from the age groups "18-20" and "21-23", but the most confident on average tend to be volunteers from the age group "24-30". 

```{r}
#| label: fig-perception-avenir-descriptive
#| fig-cap: Share of volunteers according to their confidence in their future, grouped by different variables.

plot_variables <- c("rupture", "age_categorie", "type_volontaire", "niveau_etudes")

plot_data <- combined %>%
  mutate(across(all_of(plot_variables), as.character)) %>%
  pivot_longer(cols = all_of(plot_variables), names_to = "variable", values_to = "level") %>%
  drop_na(level) |> 
  drop_na(perception_avenir) |> 
  group_by(variable, level, perception_avenir) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(variable, level) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# add correct levels for order
level_orders <- lapply(combined[plot_variables], levels)

plot_data <- plot_data %>%
  group_by(variable) %>%
  mutate(level = factor(level, levels = level_orders[[unique(variable)]])) %>%
  ungroup()

# set colours
ggplot(plot_data, aes(x = level, y = pct, fill = factor(perception_avenir))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Purples") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, y = "Share",
    fill = "Confidence in Future",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct))) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 20, hjust = 1), 
        legend.position = "top")
```

### Other factors

The only factor which we could statistically relate to confidence in one's future is `action_individuelle_societe` ("En général, pensez-vous que votre action individuelle peut contribuer à changer la société ?" Oui; Pas vraiment; Non). As shown in @fig-perception-avenir-other, volunteers who think they can contribute to society tend to have more confidence in their future than volunteers who don't. 

```{r}
#| label: fig-perception-avenir-other
#| fig-cap: Share of volunteers according to their confidence in their future, as a function of whether they think their individual action can contribute to society.

plot_variables <- c("action_individuelle_societe")

plot_data <- combined %>%
  mutate(across(all_of(plot_variables), as.character)) %>%
  pivot_longer(cols = all_of(plot_variables), names_to = "variable", values_to = "level") %>%
  drop_na(level) |> 
  drop_na(perception_avenir) |> 
  group_by(variable, level, perception_avenir) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(variable, level) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# add correct levels for order
level_orders <- lapply(combined[plot_variables], levels)

plot_data <- plot_data %>%
  group_by(variable) %>%
  mutate(level = factor(level, levels = level_orders[[unique(variable)]])) %>%
  ungroup()

# set colours
ggplot(plot_data, aes(x = level, y = pct, fill = factor(perception_avenir))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Purples") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, y = "Share",
    fill = "Confidence in Future",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct))) +
  plot_theme +
  theme(axis.text.x = element_text(angle = 40, hjust = 1), 
        legend.position = "top")
```

## Differences between key programs

@fig-perception-avenir-programs shows descriptive differences in volunteers' confidence in their future between different key programs. 

```{r}
#| label: fig-perception-avenir-programs
#| fig-cap: "Number of volunteers responding with different levels of confidence in their future, grouped by which program they were in. Note that we focus only on key programs here. All other programs appear under 'NA' in the plot."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

plot_data <- combined %>%
  drop_na(perception_avenir) |> 
  group_by(programme_cle, perception_avenir) |> 
  summarize(n = n_distinct(id_jeune)) |> 
  group_by(programme_cle) %>%
  mutate(pct = n / sum(n) * 100) |> 
  ungroup()

# absolute numbers
ggplot(plot_data, aes(x = perception_avenir, y = n, fill = factor(perception_avenir))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = n),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Purples") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, 
    y = "Number of volunteers",
    fill = "Confidence in Future",
  ) +
  scale_y_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.2))) +
  plot_theme +
  theme(axis.text.x = element_blank(), 
        legend.position = "top") +
  facet_wrap(~programme_cle)


# percentages
ggplot(plot_data, aes(x = perception_avenir, y = pct, fill = factor(perception_avenir))) +
  geom_col(position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = sprintf("%.1f%%", pct)),
    position = position_dodge(width = 0.9),
    vjust = -0.3,
    size = 3
  ) +
  scale_fill_brewer(palette = "Purples") +
  facet_wrap(~ variable, scales = "free_x") +
  labs(
    x = NULL, 
    y = "Share of volunteers",
    fill = "Satisfaction",
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1), breaks = seq(0,100, 20)) +
  coord_cartesian(ylim = c(0, max(plot_data$pct))) +
  plot_theme +
  theme(axis.text.x = element_blank(), 
        legend.position = "top") +
  facet_wrap(~programme_cle)


```

# Differences between programs {#sec-programs}

```{r}
program_count <- combined |> 
  drop_na(programme_1, programme_grouped, programme_cle) |> 
  summarize(n_programs = n_distinct(programme_1), 
            n_program_categories = n_distinct(programme_grouped),
            n_key_programs = n_distinct(programme_cle))

key_programs <- combined |>
  drop_na(programme_cle) |>
  distinct(programme_cle) |>
  pull(programme_cle) |>
  unique() |>
  paste(collapse = "; ")
```

```{r}
# define non-demographic variables to include in comparison
factors_for_program_comparison <- c(
  "satisfaction_num",
  "perception_avenir_num",
  "action_individuelle_societe",
  "comparaison_utilite_autres",
  "fierte_num",
  "confiance_en_soi_num",
  "confiance_avenir_personnel_num",
  "impact_situation_actuelle"
)
```

There are `r program_count$n_programs` different programs, which can be sorted into `r program_count$n_program_categories` different categories (see [Tables](tables.qmd)). Here, we only distinguish between `r program_count$n_key_programs` key program categories, namely: `r key_programs`. @fig-programs-overview provides an overview of the number of volunteers in the different key programs, across promotions.

In the following sections, we look at how volunteers differ demographically between the different key programs. Statistical models on program differences can be found in the Appendix (@sec-appendix-programs). Here, we only descriptively show the differences between programs for those variables that were statistically significant in these models. 
Each of the following sections focuses on one key program. For relevant variables, we show how the average in the respective key program differs compared to the average of all other key programs combined.

```{r}
#| label: fig-programs-overview
#| fig-cap: "Number of volunteers per key program, within the different promos. Note that, in the percentage plot, the percentages are relative to all volunteers from the respective promo. All volunteers who were not included in any of the key programs appear in the plot under 'NA'."
#| fig-subcap:
#|   - "(absolute numbers)"
#|   - "(percentages)" 
#| layout-ncol: 2

# ABSOLUTE NUMBERS

# Prepare data: Count unique participants per source
plot_data <- combined |> 
  group_by(promo, programme_cle) |> 
  summarise(n_participants = n_distinct(id_jeune)) |> 
  group_by(promo) |> 
  mutate(share = n_participants/sum(n_participants)) |> 
  ungroup()


# Plot attrition trend across sources
ggplot(plot_data, aes(y = programme_cle, x = n_participants, fill = programme_cle)) +
geom_col() +
  geom_text(aes(label = n_participants), hjust = -0.5, size = 3) +
  scale_fill_manual(values = program_colours) +
  scale_x_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.3))) +
  theme_minimal() +
  labs(
    x = "Program",
    y = "Number of volunteers"
  ) + 
  guides(fill = "none") + 
  plot_theme +
  facet_wrap(~promo)

# PERCENTAGES

ggplot(plot_data,
       aes(y = programme_cle, x = share, fill = programme_cle)) +
  geom_col() +
  scale_fill_manual(values = program_colours) +
  geom_text(aes(label = paste0(round(share*100, 1), "%")), hjust = -0.5, size = 3) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1), expand = expansion(mult = c(0, 0.3))) +
  labs(
    x = "Program",
    y = "Share of volunteers"
  ) +
  guides(fill = "none") + 
  plot_theme +
  facet_wrap(~promo)
```

## Solidarité Aidants {#sec-ecovolonterre}

```{r}
# Run on Solidarité Aidants data
model_results_solidarite_aidants <- run_program_models(
  data = combined,
  outcome_var = programme_solidarite_aidants_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_solidarite_aidants |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "Solidarité Aidants" program, there tend to be: more female volunteers; less refugees; fewer volunteers in the "16-17" and "24-30" age groups, but more refugees in the age groups "18-20" and "21-23"; more volunteers in rural, "QPV" and "ZRR" residential zones compared to urban ones; more educated volunteers, in particular in the "bac2" category; slightly more volunteers with a declared handicap.

```{r}
#| label: fig-solidarite-aidants-descriptive
#| fig-cap: Differences in Solidarité Aidants vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "refugie",
  "age_categorie",
  "zone_residence",
  "niveau_etudes",
  "handicap_auto_declare"
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_solidarite_aidants_binary,
  program_colour = program_colours[1]
)

```

## Cinéma & Citoyenneté {#sec-cine}

```{r}
# Run model
model_results_cine <- run_program_models(
  data = combined,
  outcome_var = programme_cine_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_cine |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "Cinéma & Citoyenneté" program, there tend to be: more male volunteers; more older volunteers; more volunteers in urbain and "ZRR" residential zones; more educated volunteers; slighly fewer handicaped volunteers; fewer volunteers who have a rupture. 

```{r}
#| label: fig-cine-descriptive
#| fig-cap: Differences in Cinéma & Citoyennetévs vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "age_categorie",
  "zone_residence",
  "niveau_etudes",
  "rupture" 
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_cine_binary,
  program_colour = program_colours[2]
)
```

## Booster {#sec-booster}

```{r}
# Run model
model_results_booster <- run_program_models(
  data = combined,
  outcome_var = programme_booster_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_booster |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "Booster" program, there tend to be: more male volunteers; considerably more younger volunteers; more volunteers in "QPV" residential zones; less educated volunteers; fewer volunteers with a declared and an adimistrative handicap; more volunteers who have a rupture.

```{r}
#| label: fig-booster-descriptive
#| fig-cap: Differences in Booster vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "age_categorie",
  "zone_residence",
  "niveau_etudes",
  "handicap_auto_declare",
  "handicap_administratif",
  "rupture" 
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_booster_binary,
  program_colour = program_colours[3]
)
```

## Ecovolonterre {#sec-ecovolonterre}

```{r}
# Run model
model_results_ecovolonterre <- run_program_models(
  data = combined,
  outcome_var = programme_ecovolonterre_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_ecovolonterre |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "Ecovolonterre" program, there tend to be: more male volunteers; considerably more refugees; more  volunteers in the "21-23" age category; more volunteers in urbain and fewer in "QPV" residential zones; more educated volunteers, in particular far more volunteers in the "bac+3" category; fewer handicapped volunteers; less volunteers who have a rupture.

```{r}
#| label: fig-ecovolonterre-descriptive
#| fig-cap: Differences in Ecovolonterre vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "refugie",
  "age_categorie",
  "zone_residence",
  "niveau_etudes",
  "rupture" 
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_ecovolonterre_binary,
  program_colour = program_colours[4]
)
```

## Médiaterre {#sec-mediaterre}

```{r}
# Run model
model_results_mediaterre <- run_program_models(
  data = combined,
  outcome_var = programme_mediaterre_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_mediaterre |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "Médiaterre" program, there tend to be: more male volunteers; more older volunteers; considerably more volunteers in "ZRR" residential zones; more educated volunteers.

```{r}
#| label: fig-mediaterre-descriptive
#| fig-cap: Differences in Ecovolonterre vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "age_categorie",
  "zone_residence",
  "niveau_etudes"
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_mediaterre_binary,
  program_colour = program_colours[5]
)
```

## ASM {#sec-asm}

```{r}
# Run model
model_results_asm <- run_program_models(
  data = combined,
  outcome_var = programme_asm_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_asm |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "ASM" program, there tend to be: more female volunteers; more  older volunteers, particular in the "21-23" age category; more volunteers in urbain and rural, considerably fewer in "QPV", and none in "ZRR" residential zones; more educated volunteers, in particular far more volunteers in the "bac+3" category; less volunteers who have a rupture.

```{r}
#| label: fig-asm-descriptive
#| fig-cap: Differences in ASM vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "age_categorie",
  "zone_residence",
  "niveau_etudes",
  "rupture" 
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_asm_binary,
  program_colour = program_colours[6]
)
```

## Solidarité Séniors {#sec-solidarite-seniors}

```{r}
# Run model
model_results_solidarite_seniors <- run_program_models(
  data = combined,
  outcome_var = programme_solidarite_seniors_binary,
  predictor_sets = demographic_variables
) 

# check
# model_results_solidarite_seniors |> 
#   filter(significant)
```

Compared to all other programs combined, among the volunteers in the "Solidarité Séniors" program, there tend to be: more female volunteers; more younger volunteers, in particular in the "18-20" age category; more volunteers in rural and in "QPV", but fewer in "ZRR" residential zones; less educated volunteers, in particular fewer volunteers in the "bac+3" category; more volunteers with a declared and with a administrative handicap; more volunteers who have a rupture.

```{r}
#| label: fig-solidarite-seniors-descriptive
#| fig-cap: Differences in Solidarité Séniors vs. other programs along demographic factors.

plot_variables <- c(
  "sexe",
  "age_categorie",
  "zone_residence",
  "niveau_etudes",
  "handicap_auto_declare",
  "handicap_administratif",
  "rupture" 
)

plot_faceted_distribution(
  data = combined,
  plot_variables = plot_variables, 
  fill_var = programme_solidarite_seniors_binary,
  program_colour = program_colours[7]
)
```

# Conclusion: Some methodological notes

This report can only provide snapshots of the extensive questionnaire data collected by Cités Unies. We aimed to cover a broad range of topics and variables in order to reveal potentially informative patterns and trends. Rather than summarizing these patterns once more, we conclude with several methodological notes that highlight opportunities to improve future impact evaluations based on the questionnaires.

First, the questionnaires could likely be shorter. Reducing the number of questions would probably decrease attrition: volunteers who respond to a very long questionnaire at q1 may be less motivated to complete another long questionnaire at q2, let alone q3. Shorter questionnaires could also improve response quality. When questionnaires are lengthy, respondents may become tired or bored and begin answering quickly and without much attention. Another option to improve response quality would be to include attention checks. That said, because volunteers are not required to take the survey, it is reasonable to assume that those who do participate are generally motivated to complete it in good faith.

Another potential improvement would be to make the questionnaires at different time points more similar. Asking the same questions across multiple waves makes it possible to track how individual volunteers change over the course of their service civique (as done in @sec-within-change).

On a minor note, it may be more informative to ask volunteers about their voting intentions rather than their past voting behavior. There are indications that some volunteers may not interpret the current question about past voting behavior as intended (@fig-vote_derniers_elections-youngest). In addition, changes in past voting behavior are inherently dependent on whether elections happen to occur for a particular cohort during their year of service civique.

Finally, I would like to reiterate a point already raised in the introduction: the questionnaire data are purely observational, and none of the associations presented in this report support specific causal interpretations. For example, in the section on demographic differences between programs (@sec-programs), programs with older volunteers also tend to have more highly educated volunteers. These observations are likely related, since older participants have simply had more time to pursue education. As another example, @fig-satisfaction-programs suggests that fewer volunteers report being "very satisfied" in key programs compared to other programs (listed in the figure under "NA"). However, key programs have attracted an increasing number of volunteers over time (@fig-programs-overview), and some of them are relatively recent. At the same time, the share of volunteers who report being "very satisfied" has declined slightly across cohorts (@fig-satisfaction). It is therefore unclear whether the observed differences are driven by the programs themselves, broader cohort trends in satisfaction, or other unobserved factors.


# Supplemental materials: Appendix 

## Voting {#sec-appendix-voting}

@fig-vote-change-model shows how different demographic variables predict changes in voting behavior. For this analysis, we look at all cohorts together. The outcome, changes in voting behavior, is binary (changed to no vs. changed to yes)^[I.e. ignoring all volunteers who did not change]. 

Plotted on the y-axis are the demographic variables, taken to predict change in voting behavior. For each categorical demographic variable, different levels are shown. For example, for the variable `age_category`, the plot shows three levels: `age_categorie18-20`, `age_categorie21-23`, `age_categorie24-30`. Each of these categories has an estimate (x-axis). These estimates are the results logistic regressions^[For each variable, a separate logistic regression has been run. The estimates from the logistic regression are exponantiated, so that the reported estimates can be interpreted as odds ratios (ORs).]. They indicate how being part of a certain category (e.g. being between 18 and 20 years old) predicts the chances of changing one's voting behavior. The estimates are "odds ratios", which can be conceived of as probabilities: An odds ratio of 1 means that this group has the same chance changing to "no" as changing to "yes". More than 1 means that this group is more likely to change to "yes". For example, an odds ratio of 2.0 means twice as likely. Less than 1 means that this group is more likely to change to "no". An odds ratio of 0.5 means half as likely. 

For each categorical variable, one category is chosen as the baseline (or reference) category. The odds ratios of all other categories are relative to that baseline. For example, for the variable `age_category`, the category "16-17" is not appearing in the graph, because it is the (invisible) baseline category. If the odds ratio for `age_category18–20` is 1.43, this means that 18 to 20 year-olds are more likely to change their vote towards "Yes", compared to 16 to 17 year-olds. More specifically, the odds of changing towards “Yes” among 18–20-year-olds are about 43% higher than among 16–17-year-olds. 

```{r}
#| label: fig-vote-change-model 
#| fig-cap: Effects of demographic factors on change in voting behavior. The outcome is binary (changed to no vs. changed to yes). The dots and their labels are the estimates of separate logistic regressions for each variable. The lines around the dots represent uncertainty in the estimates (95% confidence intervals). If these confidence intervals cross 1 (the dotted vertical line), the differences are not statistically significant, meaning we might observe them just by chance. The logarithmic scale (on the x-axis) is used so that in the visualization for the positive and negative odds ratio's to be symmetric (i.e. that 2 is as far away from 1 as is 0.5).

model_results <- model_results_voting_demographic %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  scale_x_log10() +
  labs(
    x = "Odds Ratio (log scale)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

To understand this odds ratio, take @fig-vote_derniers_elections-age: it shows that among the 16 to 17 year-olds, 285 have changed towards "No", while only 238 have changed towards "Yes". That makes the odds of changing towards yes for the youngest age group, chosen as the baseline category in the model, `238:285 = ~0.84`. By contrast, among the 18 to 20 year-olds, the odds were `784:658 = ~1.19`. The odds ratio for the 18 to 20 year-olds compares their odds to those of the 16 to 17 year-olds `odds ratio = 1.19:0.84 = 1.4`. @fig-vote_derniers_elections-age also allows to compare the odds ratios to the percentages for the different age groups. Just like for absolute numbers, the odds ratios can be calculated from the percentages. For example, among the 16 to 17 year-olds, 13.8% have changed towards "No", while 11.5% have changed towards "Yes", making the odds of changing to "Yes", as above, `0.115:0.138 = ~0.83`^[Note that in this illustration, we obtain a slightly different value compared to 0.84 above due to rounding imprecision.]. Among the 18 to 20 year-olds, accordingly, the odds were, as above, `0.174:0.146 = ~1.19`, resulting in the same odds ratio of ~1.4. 

## Individual action for society {#sec-appendix-individual-action}

@fig-action-change-model shows how different demographic variables predict changes in perceptions of one's individual contribution to society. For this analysis, we look at all cohorts together. The outcome, changes in perceptions of one's individual contribution to society, collapsed into a binary measure ("vers Non/Pas vraiment" vs. "vers Oui"). The plots can be interpreted as described above in the section on voting behavior (@sec-appendix-voting). An odds ratio of 1 means that a certain group has the same chance of changing to "Non/Pas vraiment" as changing to "Oui", compared to the baseline group. More than 1 means that this group is more likely to change to "Oui". Less than 1 means that this group is more likely to change to "Non/Pas vraiment". 

```{r}
#| label: fig-action-change-model 
#| fig-cap: Effects of demographic factors on change in perceptions of how one's individual actions can contribute to society. The outcome is binary ("vers Non/Pas vraiment" vs. "vers Oui"). The dots and their labels are the estimates of separate logistic regressions for each variable. The lines around the dots represent uncertainty in the estimates (95% confidence intervals). If these confidence intervals cross 1 (the dotted vertical line), the differences are not statistically significant, meaning we might observe them just by chance. The logarithmic scale (on the x-axis) is used so that in the visualization for the positive and negative odds ratio's to be symmetric (i.e. that 2 is as far away from 1 as is 0.5).

model_results <- model_results_action %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  scale_x_log10() +
  scale_y_discrete(expand = expansion(mult = c(0, 0.1))) +
  labs(
    x = "Odds Ratio (log scale)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal() 
```

## Rupture {#sec-appendix-rupture}

For analyses on what predicts ruptures, we focus only on volunteers who ended their contract early for apparently *negative* reasons. We used a binary outcome measure (no negative rupture vs. negative rupture) and ran separate logistic regressions with demographic variables (@fig-rupture-model-demographic) a selection of other variables (@fig-rupture-model-other). 

An odds ratio of 1 means that a certain group has the same chance of having a rupture, compared to the baseline group. More than 1 means that this group is more likely to have a rupture. Less than 1 means that this group is less likely to have a rupture.

### Demographics

```{r}
#| label: fig-rupture-model-demographic 
#| fig-cap: Effects of demographic factors on negative ruptures. The outcome is binary (no negative rupture vs. negative rupture). The dots and their labels are the estimates of separate logistic regressions for each variable. The lines around the dots represent uncertainty in the estimates (95% confidence intervals). If these confidence intervals cross 1 (the dotted vertical line), the differences are not statistically significant, meaning we might observe them just by chance. The logarithmic scale (on the x-axis) is used so that in the visualization for the positive and negative odds ratio's to be symmetric (i.e. that 2 is as far away from 1 as is 0.5).


model_results <- model_results_rupture_demographic_variables %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  scale_x_log10() +
  labs(
    x = "Odds Ratio (log scale)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Other factors

@fig-rupture-model-other shows how the two relevant variables predict rupture. The variable `action_individuelle_societe` is categorical and can be interpreted like the demographic variables in the previous results: one category serves as the reference (omitted) category, and the reported odds ratios represent how the odds of rupture differ relative to that reference. The variable `perception_avenir_num` is numeric (on a scale from 1 to 4)^[Note `_num` in the variable name stands for 'numeric': we use a numeric version, from "Pas du tout confiant(e)" = 1, to "Très confiant(e)" = 4.]. For numeric predictors, the estimated odds ratio indicates how much the odds of rupture change, on average, for each one-unit increase on the scale. For example, the resulting estimate of 1.14 in @fig-rupture-model-other for `perception_avenir_num` means that, on average, a one point increase in confidence about one's future corresponds to 14% higher odds of rupture. In other words, the results suggest that, the more confident volunteers are about their future, the more likely they are to have a rupture, on average. @fig-perception-avenir-descriptive suggests that this average effect is mostly driven by those volunteers who are "very confident". 

```{r}
#| label: tbl-rupture-candidates
#| tbl-cap: Candidate variables to evaluating their association with rupture.

factors_for_rupture <- c("perception_avenir", 
                         "projet_avenir_concret", "action_individuelle_societe", 
                         "comparaison_utilite_autres", "fierte",
                         "confiance_en_soi", "confiance_avenir_personnel",
                         "impact_situation_actuelle", "integration")


combined |> 
  pivot_longer(all_of(factors_for_rupture), 
               names_to = "variable", 
               values_to = "value") |> 
  drop_na(value) |> 
  distinct(variable, source) |> 
  kbl()
```

```{r}
#| label: fig-rupture-model-other 
#| fig-cap: Effects of demographic factors on negative ruptures. The outcome is binary (no negative rupture vs. negative rupture). The dots and their labels are the estimates of separate logistic regressions for each variable. The lines around the dots represent uncertainty in the estimates (95% confidence intervals). If these confidence intervals cross 1 (the dotted vertical line), the differences are not statistically significant, meaning we might observe them just by chance. The logarithmic scale (on the x-axis) is used so that in the visualization for the positive and negative odds ratio's to be symmetric (i.e. that 2 is as far away from 1 as is 0.5).


model_results <- model_results_rupture_other_variables %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  scale_x_log10() +
  labs(
    x = "Odds Ratio (log scale)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

## Satisfaction {#sec-appendix-satisfaction}

For analyses on what predicts satisfaction, we treat satisfaction as a continuous variable (from 1, "Pas du tout satisfaisante", to 4, "Très satisfaisante"). Because this outcome is continuous, we run linear regressions (and not logistic regressions, as before) on it. The results for demographic variables are displayed in @fig-rupture-model-demographic, for other variables in @fig-rupture-model-other. 

### Demographic factors

The results of the linear regressions are more straightforward to interpret. For categorical predictors (i.e., all demographic variables), the estimate is simply the average difference between the shown category and the (again, omitted) baseline category. For example, as shown with the variable `refugie`, refugees tend to be, on average, more satisfied by 0.2 points (on the 4-point satisfaction scale) than non-refugees.

```{r}
#| label: fig-satisfaction-demographic-variables 
#| fig-cap: Effects of demographic factors on satisfaction.


model_results <- model_results_satisfaction_demographic_variables %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate (on a scale from 1 to 4)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Other factors

Some of the non-demographic variables are numeric^[All variable names that end with `_num` have been treated as numeric for the present analyses], for example, `confiance_en_soi_num` (Pensez-vous avoir gagné en confiance en vous durant votre Service Civique ? 1 - Non, pas du tout; 2 - Non, pas vraiment; 3 - Oui, assez; 4 - Oui, beaucoup). 

For numeric variables, the estimates represent how much satisfaction increases or decreases with a one-unit increase in the predictor variable. For example, for `confiance_en_soi_num`, the estimate is 0.33. `confiance_en_soi_num` takes values on a 4-point scale (from 1, "Non, pas du tout", to 4, "Oui, beaucoup"). This means that, on average, a one unit increase in `confiance_en_soi_num` is associated with a 0.33 point in crease in satisfaction (4-point scale). 

```{r}
#| label: fig-satisfaction-other-variables 
#| fig-cap: Effects of other, non-demographic factors on satisfaction.

model_results <- model_results_satisfaction_other_variables %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate (on a scale from 1 to 4)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

## Confidence in one's future {#sec-appendix-perception-avenir}

As for satisfaction above, for analyses on what predicts confident in their future, we treat it as a continuous variable (1, "Pas du tout confiant.e", to 4, "Très confiant.e"). As for satisfaction, because this outcome is continuous, we run linear regressions on it. The results for demographic variables are displayed in @fig-perception-avenir-demographic-variables, for other variables in @fig-perception-avenir-other-variables. 

### Demographic factors

```{r}
#| label: fig-perception-avenir-demographic-variables 
#| fig-cap: Effects of demographic factors on confidence in one's future.


model_results <- model_results_perception_demographic_variables %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate (on a scale from 1 to 4)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Other factors

The only other variable we consider here is `action_individuelle_societe` ("En général, pensez-vous que votre action individuelle peut contribuer à changer la société ?" Oui; Pas vraiment; Non). @fig-perception-avenir-other-variables shows that volunteers who do not think that they can individually contribute to changing society (the omitted baseline category) are the least confident about their future: both volunteers who are slightly less pessimistic (having answered "Not really") and volunteers think they can contribute (having answered "Yes) are more likely to be confident, with those having answered "Yes" being most likely.

```{r}
#| label: fig-perception-avenir-other-variables 
#| fig-cap: Effects of other, non-demographic factors on confidence in one's future.

model_results <- model_results_perception_other_variables %>%
  mutate(
    label = round(estimate, digits = 2)
  )

ggplot(model_results, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate (on a scale from 1 to 4)",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

## Differences between programs {#sec-appendix-programs}

The following sections show, for each key program category, how the volunteers of that program differ compared to all the other key program categories combined, in terms of demographic variables. The results are expressed in odds ratios, which can be interpreted as above: If the odds ratio is *larger than 1*, then a certain group is *more likely* to be represented in the target key program category, compared to all other programs. For example, in @fig-solidarite-aidants-demographic-variables, there are more 18 to 20 year-olds (odds ratio = 1.18, compared to the omitted baseline, 16 to 17 year-olds) in the "Solidarité Aidants" program, than in the other programs, on average. If the odds ratio *is below 1*, this means that a certain group is *less likely* to be represented in the target key program category. For example, refugees are considerably less likely (odds ratio = 0.17, compared to the omitted baseline, non-refugees) to be in the "Solidarité Aidants" program, compared to all other programs, on average. 

### Solidarité Aidants 

```{r}
#| label: fig-solidarite-aidants-demographic-variables 
#| fig-cap: Differences in Solidarité Aidants vs. other programs along demographic factors.

ggplot(model_results_solidarite_aidants , aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Cinéma & Citoyenneté

```{r}
#| label: fig-cine-demographic-variables 
#| fig-cap: Differences in Cinéma & Citoyennetévs vs. other programs along demographic factors

ggplot(model_results_cine, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Booster

```{r}
#| label: fig-booster-demographic-variables 
#| fig-cap: Differences in Booster vs. other programs along demographic factors.

ggplot(model_results_booster, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Ecovolonterre

```{r}
#| label: fig-ecovolonterre-demographic-variables 
#| fig-cap: Differences in Ecovolonterre vs. other programs along demographic factors.

ggplot(model_results_ecovolonterre, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Médiaterre

```{r}
#| label: fig-mediaterre-demographic-variables 
#| fig-cap: Differences in Médiaterre vs. other programs along demographic factors.

ggplot(model_results_mediaterre , aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### ASM 

```{r}
#| label: fig-asm-demographic-variables 
#| fig-cap: Differences in ASM vs. other programs along demographic factors.

ggplot(model_results_asm, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```

### Solidarité Séniors

```{r}
#| label: fig-solidarite-seniors-demographic-variables 
#| fig-cap: Differences in Solidarité Séniors vs. other programs along demographic factors.

ggplot(model_results_solidarite_seniors, aes(x = estimate, y = fct_rev(term), color = variable)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray50") +
  geom_point() +
  geom_text(
    aes(label = label),
    vjust = -0.4,  # just slightly to the right of the point
    size = 3,
    check_overlap = TRUE  # avoid overlapping labels
  ) +
  scale_x_log10() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  scale_color_viridis_d(option = "plasma", end = 0.8) +
  labs(
    x = "Estimate",
    y = NULL
  ) +
  guides(color="none") +
  theme_minimal()
```




